---
title: "Quant_QC_DWI.Rmd"
author: "nforde"
date: "October 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prelude
This report should give you details of the DWI Quantitative QC for your datasets.  
```{r load, echo=FALSE, warnings=FALSE, message=FALSE}
args <- commandArgs(TRUE)

if (length(args)==0) {
  stop("At least one argument must be supplied (csv)", call.=FALSE)
}
# else if (length(args)==1) { #Ccan use something like this so the SD can be changeable
#   # default SD size is 2
#   args[2] = 2
# }

metric_file = args[1]
#SD_size = args[2]
outdir <- dirname(metric_file)

#loading necessary libraries
library("tidyverse", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")
library("cowplot", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")
library("ggpubr", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")
library("gridExtra", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")
library("grid", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")
library("rmarkdown", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")
library("kableExtra", lib.loc="/mnt/tigrlab/quarantine/R/3.4.3/build2/lib/R/library")

```
# Plots of each metric
```{r plots, echo=FALSE}
## set paths & load csv

df <- read.csv(metric_file, header=TRUE) 

#bvals can vary across studies and there are multiple for multishell data. This will make a list of the cnr variables you need 
avg_cnr <- df %>% select(starts_with("avg_cnr")) %>% names()

plot_cnr <- function(cnr, df) {
  df2 <- df %>% filter(get(cnr) < 9999)
  plot <- ggplot(df2, aes_string(x = cnr)) +
    geom_density(size = 2, aes(color = "#440154FF"), fill="lightblue") +
    geom_vline(data = df2, aes(xintercept = mean(get(cnr)), color = "lightblue"), linetype = "dashed", size = 1) +
    labs(title = paste0("Distribution of ", cnr," CNR Values"), x = paste0(cnr, " CNR Values"), y = "Density") +
    theme_bw() + theme(legend.position="none")
  return(plot)
}

density_plot_cnr <- lapply(avg_cnr, function(f) plot_cnr(f, df)) # this should create a list of 1 or more plots depending on the number of bvals

density_plot_snr <- function(df) {
  df2 <- df %>% filter(avg_snr_0 < 9999)
  plot <- ggplot(df2, aes(x = avg_snr_0)) +
    geom_density(size = 2, aes(color = "#440154FF"), fill = "lightblue") +
    geom_vline(data = df2, aes(xintercept = mean(avg_snr_0), color = "lightblue"), linetype = "dashed", size = 1) +
    labs(title = "Distribution of SNR Values", x = "SNR Values", y = "Density") +
    theme_bw() + theme(legend.position="none")
  return(plot)
}

density_plot_rel_mot <- function(df) {
  plot <- ggplot(df, aes(x = qc_mot_rel)) +
    geom_density(size = 2, aes(color = "#440154FF"), fill = "lightblue") +
    geom_vline(data = df, aes(xintercept = mean(qc_mot_rel), color = "lightblue"), linetype = "dashed", size = 1) +
    labs(title = "Distribution of Relative Motion Values", x = "Relative Motion Values", y = "Density") +
    theme_bw() + theme(legend.position="none")
  return(plot)
}

density_plot_outliers <- function(df) {
  plot <- ggplot(df, aes(x = qc_outliers_tot)) +
    geom_density(size = 2, aes(color = "#440154FF"), fill = "lightblue") +
    geom_vline(data = df, aes(xintercept = mean(qc_outliers_tot), color = "lightblue"), linetype = "dashed", size = 1) +
    labs(title = "Distribution of Outlier Values", x = "Outlier Values", y = "Density") +
    theme_bw() + theme(legend.position="none")
  return(plot)
}

plot_grid(density_plot_outliers(df), density_plot_rel_mot(df), density_plot_snr(df) + rremove("x.text"), 
          labels = c("A", "B", "C"),
          ncol = 1, nrow = 3)

plot_grid(plotlist=density_plot_cnr, ncol = 1)  

```
  
# Outliers  
Outliers are calculated as more than 2 standard deviations from the group average for that metric. In some cases we only exclude those more than 2 SD in the _bad_ direction. i.e. a subject with extremely low motion may be more than 2 SD from the mean but we will not exclude for that.   
```{r outliers, echo=FALSE}

#Excluding based on bad CNR
# bad_cnr_1500 <- df %>% filter(avg_cnr_1500 < 9999) %>%
#   mutate(cnr_std = scale(avg_cnr_1500)) %>% #get SD 
#   filter(cnr_std < -2 | cnr_std > 2) #filter subjects who had CNR that was > and < than 2 SD's from the mean
# 
# #Note: if its multishell then you will have different average CNRs for every B values, and in that case, be sure to check all of them
# #In the case of multishell acquisitions, use this if loop and feel free to change the B values to suit your study specific acquisitions
# 
# # bad_cnr_1600 <- df %>%
# #   mutate(cnr_std_1600 = scale(avg_cnr_1600)) %>% 
# #   filter(cnr_std_1600 < -2 | cnr_std_1600 > 2)
# # 
# # bad_cnr_2600 <- df %>%
# #   mutate(cnr_std_2600 = scale(avg_cnr_2600)) %>% 
# #   filter(cnr_std_2600 < -2 | cnr_std_2600 > 2)
# 
# 
# #Excluding based on bad SNR
# bad_snr <- df %>%
#   mutate(snr_std = scale(avg_snr_0)) %>%
#   filter(snr_std < -2 | snr_std > 2)
# 
# #Excluding based on too much relative motion
# bad_rel_mot <- df %>%
#   mutate(rel_mot_std = scale(qc_mot_rel)) %>%
#   filter(rel_mot_std > 2) #for motion, we're only concerned about participants with too much motion
# 
# #Excluding based on too many outliers
# bad_outliers <- df %>%
#   mutate(outliers_std = scale(qc_outliers_tot)) %>%
#   filter(outliers_std > 2)  #for outliers, we're only concerned about participants with too many outliers
# 
# #Creating a dataframe all the subjects that failed the Quantitative QC and their problematic values
# 
# Failed_Quant_QC <- rbind.fill(bad_cnr1500, bad_snr, bad_rel_mot, bad_outliers)
# 
# Failed_Quant_QC_DF <- Failed_Quant_QC[,c("subject_id", "avg_cnr_1000", "avg_cnr_1600", "avg_cnr_2600", "avg_snr_0", "qc_mot_rel", "qc_outliers_tot")]

# out.file <- paste(outdir, "Quant_QC_2SD.csv", sep="/")
# write.csv(Failed_Quant_QC_DF, out.file, row.names = T)
# 
# DWI.Metrics.Table <- as.data.table(Failed_Quant_QC_DF, keep.rownames = TRUE)
# DWI.Metrics.Table %>%
#   kable() %>%
#   kable_styling()



```